\documentclass[12pt,ifthen]{article}
\usepackage{url}
\usepackage{comment}

\newif{\ifshowsoln}
% \showsolntrue

\newcommand{\und}{\_\_\_\_\_\_\_\_\_}
\newcommand{\Z}{\mathbb{Z}}
\usepackage{amsmath}
\usepackage{amssymb}  % for \nmid

\begin{document}

{\textbf Lecture Notes 2/2/2026

CMSC250H

\vspace{1 cm}
}
\textbf{Circuits}

\begin{itemize}
   \item Gates are expensive, so it's in our best interest to make them cheaper and more simplified
   \begin{itemize}
    \item For example, use DeMorgan's Law to simplify gates
   \end{itemize}
   \item We want to build circuits that add arbitrarily large binary numbers
   \begin{itemize}
    \item e.g 10011001 + 00110011 = 11001100
   \end{itemize} 
   \item A half-adder is a circuit that adds two bits together!
   \begin{itemize}
    \item e.g X + Y = CS, where CS is the carry bit
    \item We can treat the half-adder as an abstraction since we already know how to build it.
   \end{itemize}
   \item Exclusive or(XOR): one or other, but not both
   \item $\oplus$ is the XOR gate
   \item A Full-Adder can compute the sum of two 2-digit binary numbers.
   \begin{itemize}
    \item We need the ability to add 3 digits because a C might carry over from the largest digit addition
    \item This circuit that can add 3 bits is a full-adder.
    \item These can also be abstracted.
    \item Finally, we can connect the Half-Adder and its new inputs to the Full-Adder to produce the final circuit.
   \end{itemize}
   \item To construct an n-bit adder, we have n-1 full adders
   \item There are circuits for addition, multiplication, division, and a primality test, along with more.
\end{itemize}
\vspace{0.9cm}
\textbf{Intermediary Truth Values} 
\begin{itemize}
    \item In math, statements are usually true or false (1 or 0)
    \item But we want truth values between 0 and 1. Like the chance that it rains tomorrow, which is a probability.
    \item $\wedge$ can be defined as xy, and $\vee$ can be defined as x+y-xy. $\neg$ can be defined as 1-x or 1-y.
    \item Another way to define $\wedge$ is min(x,y), while $\vee$ can be defined as max(x,y).
\end{itemize}

\textbf{To-Do After Class: }
\begin{itemize}
    \item Look at the actual construction of half-adders and full-adders.
\end{itemize}



\end{document}